\documentclass[letterpaper,11pt]{memoir}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{mathpazo} % Palatino fonts for mathematics
\usepackage{tgpagella} % improved Palatino font for text
\usepackage{microtype} % better justification
\usepackage{amsmath}
\usepackage{cmap} % make PDFs searchable

% hyperref turns internal references to hyperlinks, allows \url{}
% also allows setting PDF meta information
\usepackage[pdftex,
  pdfauthor={Alex Reinhart},
  pdftitle={An Integrated System for Gamma-Ray Spectral Mapping and Anomaly
    Detection}]{hyperref} 

%% Memoir options
\chapterstyle{ell}
% Put section numbers in left margin
\setsecindent{-33pt}
% Include subsections in TOC
\setcounter{tocdepth}{3}
\setsecnumdepth{subsection}
% Caption fonts
\captionnamefont{\small\itshape}
\captiontitlefont{\small}
\captionstyle[\centering]{}

% Signature line
\newcommand{\sigline}{\noindent
\makebox[2.5in]{\hrulefill}\\}
% Date line -- shorter
\newcommand{\dateline}{\noindent
\makebox[1.5in]{\hrulefill}\\}

\begin{document}
\pagenumbering{roman}

% Title page
\begin{titlingpage}
\begin{center}
\begin{Spacing}{2}
{\LARGE
  An Integrated System for Gamma-Ray Spectral Mapping and Anomaly Detection
}
\end{Spacing}
\end{center}

\begin{center}
  Presented by Alex Reinhart\\
  Applied Research Laboratories\\
  The University of Texas at Austin
\end{center}

\begin{center}
In partial fulfillment of the requirements for graduation with the\\
Dean’s Scholars Honors Degree in Physics
\end{center}

\vskip 1in

\begin{minipage}[t]{0.5\textwidth}
\sigline 
Dr.~Alex Athey\\
Supervisor
\end{minipage}\hskip 1in
\begin{minipage}[t]{0.4\textwidth}
\dateline
Date
\end{minipage}

\vskip 0.85in

\begin{minipage}[t]{0.5\textwidth}
\sigline
Prof.~Roy Schwitters\\
Co-Supervising Professor
\end{minipage}\hskip 1in
\begin{minipage}[t]{0.4\textwidth}
\dateline
Date
\end{minipage}

\vskip 0.85in

\begin{minipage}[t]{0.5\textwidth}
\sigline
Prof.~Greg Sitz\\
Honors Adviser in Physics
\end{minipage}\hskip 1in
\begin{minipage}[t]{0.4\textwidth}
\dateline
Date
\end{minipage}
\end{titlingpage}

\cleardoublepage

\chapter*{Acknowledgments}

This work would not be possible without the support of my supervisor, Dr.~Alex
Athey, along with his colleagues at Applied Research Laboratories. Many helpful
suggestions were provided by Dr.~Steve Biegalski and Tracy Tipping of the
Nuclear Engineering Teaching Laboratories, along with experimental
support. Dr.~James Scott gave much valuable statistical advice. Bridgeport
Instruments provided helpful guidance in using their scintillator detectors. I
benefitted from the wisdom of my thesis advisor, Dr.~Roy Schwitters, and his
many suggestions. Scott Pennington lent his expertise and worked to ensure the
cooperation of UT Athletics, making our football stadium data collection
possible.

This research has been funded in part by the Office of the Deputy Assistant
Secretary of Defense for Nuclear Matters.

Finally, I must thank Rey Gomez, Patrick Vetter, Franziska Klingberg, and Todd
Hay, whose legs suffered in the name of science.

\cleardoublepage

\chapter*{Abstract}
For security, environmental, and regulatory purposes it is useful to monitor
wide areas for unexpected changes in radioactivity.  Background radiation from
naturally radioactive materials presents the largest challenge to detection
sensitivity.  Most commonly, a single background spectrum is recorded and all
subsequent spectra are compared to this static background. We have developed a
temporal anomaly detection algorithm which uses multiple passes across an area
to build a spatial map of background spectra, allowing increased sensitivity to
anomalies. By comparing spectral shape rather than count rate we increase
sensitivity and limit the influence of the background-dominated low-energy
region. To demonstrate this technique, we performed source injection
simulations, simple source detection tests, and blind tests at University of
Texas football games. Applications in wide-area monitoring through detectors on
vehicles, such as buses, are explored. We also probe the feasibility of using
kriging methods from geostatistics to provide more accurate anomaly mapping.

\cleardoublepage

\tableofcontents

\cleardoublepage

\pagenumbering{arabic}

\input{introduction.tex}

\input{methods.tex}

\input{experiments.tex}

\input{kriging.tex}

\input{conclusions.tex}

\bibliographystyle{ieeetr}
\bibliography{algorithms}

\end{document}